---
title: 'Statistical Learning: Chapter 2'
author: "Phuong Dong Le"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning = FALSE} 
library(ggplot2)
library(tidyverse)
library(gridExtra)
library(kableExtra)
library(GGally)
library(table1)
```




## Exercise 8. 

This relates to the `College` data set, containing a number of variables for 777 different universities and colleges in the US. 


**(a). Use the `read.csv()` to read the data into `R`, call the loaded data `college`.**

```{r}
## set the working directory: 
setwd(dir = "~/Desktop/Statistical Learning/dataset")

### read the dataset college
college = read.csv(file = "college.csv")
```


**(b). Look at the data using the `fix()` function. Try the following commands:**

```{r}
##rownames(college) = college[,1]
##fix(college)


### college = college[,-1]
### fix(college)
```


**(c)**

**(i). Use the `summary()` to produce a numerical summary of the variables in the data set**


```{r}
TableSummary = summary(college)
print(TableSummary)
```


**(ii). Use the `pair()` function to produce the scatterplot matrix of first ten columns or variables of the data**


```{r}
PairData = college[,1:10]
pairs(x = PairData)
```


**(iii). Make a boxplot of side-by-side of Outstate versus Private**

```{r}
Boxplot1 = ggplot(data = college, mapping = aes(x = Private, y = Outstate)) + 
  geom_boxplot() + 
  ggtitle(label = "Boxplot of Outstate versus Private") + 
  stat_summary(fun.y = mean, geom = "point", shape = 23, size = 4 ) + 
  xlab(label = "Private") + 
  ylab(label = "Outstate") + 
  theme_bw()

print(Boxplot1)
```


**(iv). Create a qualitative variable `Elite`: binning the `Top10perc` variable: divide the universities into 2 groups based on prop. of students comming from the top 10% of high school exceeds 50%**

```{r}
Elite = rep("No", nrow(college))
Elite[college$Top10perc > 50] = "Yes"
Elite = as.factor(Elite)

### bind the dataframe
college = data.frame(college, Elite)

summary(college)

Boxplot2 = ggplot(data = college, mapping = aes(x = Private, y = Outstate)) + 
  geom_boxplot() + 
  ggtitle(label = "Boxplot of Outstate versus Private") + 
  stat_summary(fun.y = mean, geom = "point", shape = 23, size = 4 ) + 
  xlab(label = "Private") + 
  ylab(label = "Outstate") + 
  theme_bw()

print(Boxplot2)
```



**(v), Make histograms for a few quantitative variables**

```{r, message=FALSE, warning=FALSE}


hist1 = ggplot(data = college, mapping = aes(x = Outstate)) + 
  geom_histogram(fill = "blue") + 
  ggtitle(label = "Histogram of Outstate") + 
  theme_bw()

hist2 = ggplot(data = college, mapping = aes(x = Accept)) + 
  geom_histogram(fill = "navy") + 
  ggtitle(label = "Histogram of Accepted Applicants") + 
  theme_bw()

hist3 = ggplot(data = college, mapping = aes(x = Enroll)) + 
  geom_histogram(fill = "orange") + 
  ggtitle(label = "Histogram of New students enrolled") + 
  theme_bw()

hist4 = ggplot(data = college, mapping = aes(x = Apps)) + 
  geom_histogram(fill = "red") + 
  ggtitle(label = "Histogram of Applications") + 
  theme_bw()


grid.arrange(hist1, hist2, ncol = 2)

grid.arrange(hist3, hist4, ncol = 2)
```







## Exercise 9: 


**(a). Find the predictors: quantitative, qualitative**


```{r, message=FALSE, warning=FALSE}
## set the working directory: 
setwd(dir = "~/Desktop/Statistical Learning/dataset")

### read the datafile: 

auto = read.csv(file = "Auto.csv")
### print out the structure: 
str(auto)

```


* The quantitative predictors: `mpg`, `displacement`, `weight`, `acceleration`. 

* The qualitative predictors: `cylinders`, `horsepower`, `year`, `origin`, `name`. 



**(b). Find the range of each quantitative predictor**


```{r}
## range of mpg: 

range_mpg = range(auto$mpg)
range_displacement = range(auto$displacement)
range_weight = range(auto$weight)
range_acc = range(auto$acceleration)


print(list(
  range_mpg = range_mpg, 
  range_displacement = range_displacement, 
  range_weight = range_weight, 
  range_acc = range_acc 
))
```



**(c). Find the mean and standard deviation of each quantitative predictor**


```{r}
op_auto  = auto[,c("mpg", "displacement", "weight", "acceleration")]

print(list(
  mean_mpg = mean(op_auto$mpg), 
  mean_dis = mean(op_auto$displacement), 
  mean_weight = mean(op_auto$weight), 
  mean_acc = mean(op_auto$acceleration)
))


print(list(
  sd_mpg = sd(op_auto$mpg), 
  sd_dis = sd(op_auto$displacement), 
  sd_weight = sd(op_auto$weight), 
  sd_acc = sd(op_auto$acceleration)
))
```






**(d). Remove the 10th through 85th obsevation. Find the range, mean, and std. of each predictor in the subset of the data that remains**


```{r}
sub_auto = auto[-c(10:85),]

range_mpg = range(sub_auto$mpg)
range_displacement = range(sub_auto$displacement)
range_weight = range(sub_auto$weight)
range_acc = range(sub_auto$acceleration)


print(list(
  range_mpg = range_mpg, 
  range_displacement = range_displacement, 
  range_weight = range_weight, 
  range_acc = range_acc 
))

op_auto  = sub_auto[,c("mpg", "displacement", "weight", "acceleration")]

print(list(
  mean_mpg = mean(op_auto$mpg), 
  mean_dis = mean(op_auto$displacement), 
  mean_weight = mean(op_auto$weight), 
  mean_acc = mean(op_auto$acceleration)
))


print(list(
  sd_mpg = sd(op_auto$mpg), 
  sd_dis = sd(op_auto$displacement), 
  sd_weight = sd(op_auto$weight), 
  sd_acc = sd(op_auto$acceleration)
))

```


**(e). Make some plots highlighting the relationships among the predictors using the full data set**


```{r}
plot1  = ggplot(data = auto, aes(x = weight, y = mpg)) + 
  geom_point() + 
  ggtitle(label = "Scatterplot of mpg vs. weight") + theme_bw()

plot2  = ggplot(data = auto, aes(x = displacement, y = mpg)) + 
  geom_point() + 
  ggtitle(label = "Scatterplot of mpg vs. displacement") + theme_bw()

plot3 =  ggplot(data = auto, aes(x = acceleration, y = mpg)) + 
  geom_point() + 
  ggtitle(label = "Scatterplot of mpg vs. acceleration") + theme_bw()

grid.arrange(plot1,plot2,plot3, ncol = 2)
```


**(f). We want to predict gas mileage (mpg) on the basis of other variables**


* Based on the scatterplots above, we can predict `mpg` using the variables: `weight` and `displacement`. There is a linear negative trend between these variables. 



## Exercise 10: 


**(a). This exercise involves the `Boston` housing data set**


```{r}
library(MASS)

## Read about the dataset: 
## ?Boston

dim(Boston)
names(Boston)
head(Boston)
```

* There are 506 rows and 14 columns in data set. `crim` represent the per capita crime rate by town. `zn` is the prop. of residential land zoned for lots over 25,000 sqtft. `indus` is prop. of non-retail business acres per town. `chas` is Charles River dummy variable. `ncox` is nitrogent oxides concerntration. `rm` is average number of rooms per dwelling. `age` is prop. of owner-occupited units built prior to 1940. `dis` is weighted mean of distances to five Boston emp. centres.`rad` is index of acceibility of radial highways. `tax` is full prop. tax rate. `pratio` is pupil-teacher ratio by town. `black` prop. of blacks by town.`lstat` lower status of pop. percent. `medv` median value of owner occupied homes. 


**(b). Make some pairwise scatterplots**

```{r}
str(Boston)
Boston$chas <- as.numeric(Boston$chas)
Boston$rad <- as.numeric(Boston$rad)
pairs(x = Boston)
```

* Not much can be discerned other than the fact that some variables appear to be correlated. A correlation matrix would be more helpful and question-c gives us the opportunity to make one.

**(c). Make a correlation matrix to find association with the per capita crime rate**

```{r}

round(
  cor(x = Boston$crim, y = Boston[,c("zn", "indus","chas","nox", "rm", "age", "dis", "rad",
                                     "tax", "ptratio","black", "lstat","medv")]), 
  digits = 3)
```

* Based on the correlation coefficients and their corresponding p-values, there is indeed an association between the per capita crime rate (crim) and the other predictors.


**Part (d)**

```{r}
summary(Boston$crim)
summary(Boston$tax)
summary(Boston$ptratio)

## Histogram: 
Hist1 = ggplot(data = Boston, aes(x = crim)) + 
  ggtitle(label = "Histogram of Crime Rate") + 
  ylab(label = "Number of Suburbs") + geom_histogram(binwidth = 5)

Hist2 = ggplot(data = Boston, aes(x = tax)) + 
  ggtitle(label = "Histogram of Tax Rate ") + 
  ylab(label = "Number of Suburbs") + geom_histogram(binwidth = 5)

Hist3 = ggplot(data = Boston, aes(x = ptratio)) + 
  ggtitle(label = "Pupil-Teacher Ratio ") + 
  ylab(label = "Number of Suburbs") + geom_histogram()

grid.arrange(Hist1, Hist2, Hist3, ncol = 2)


```


* Considering that the median and maximum crime rate values are respectively about 0.26% and 89%, there are indeed some neighborhoods where the crime rate is alarmingly high

```{r}
selection <- subset( Boston, crim > 10)
nrow(selection)/ nrow(Boston)
```

* There is 11% of the neighborhood’s have crime rates above 10%

```{r}
selection <- subset( Boston, crim > 50)
nrow(selection)/ nrow(Boston)
```

* There is 0.8% of the neighborhoods have crim rates above 50%. 

* Based on the histogram of the Tax rates, they are few neighborhoods where rates are relative higher. The median and average tax amount are $330 and $408.20 ( per Full-value property-tax rate per $10,000) respectively.

```{r}
selection <- subset( Boston, tax< 600)
nrow(selection)/ nrow(Boston)
```

* 73% of the neighborhood pay under $600

```{r}
selection <- subset( Boston, tax> 600) 
nrow(selection)/ nrow(Boston)
```

* There is around 27% of the neighborhood pay over $600



**(e). The number of surbibs in this data set bound to the Charles River**

```{r}
nrow(subset(Boston, chas == 1))
```

* There are 35 such surburbs bound to the Charles River. 

**(f). The median pupil-teacher ratio among the towns**

```{r}
median(Boston$ptratio)
```

* The median is 19.05 


**(g). Find the least median surburb? values of other predictors for that surburb, and compare to the overall ranges for those predictors.**


```{r}

selection <- Boston[order(Boston$medv),]
selection[1,]

```


* Suburb #399 with a median value of $5000. We can use the following summary information to answer part-2 of this question

* Crime is very high compared to median and average rates of all Boston neighborhoods. 

* No residential land zoned for lots over 25,000 sq.ft. This applies to more than half of the neighborhoods in Boston * Proportion of non-retail business acres per town is very high compared to most suburbs. 


* This suburd is not one of the suburbs that bound the Charles river. 

* Nitrogen oxides concentration (parts per 10 million) is one of the highest. 

* Average number of rooms per dwelling is one of the lowest 


* Highest proportion of owner proportion of owner-occupied units built prior to 1940. 

* One of the lowest weighted mean of distances to five Boston employment centres. * Highest index of accessibility to radial highways. 

* One of the highest full-value property-tax rate per $10,000. 

* One of the highest pupil-teacher ratio by town * Highest value for 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town. 

* One of the highest lower status of the population (percent) 

* Lowest median value of owner-occupied homes in $1000s.

Based on the list above, suburb 399 can be classified as one of the least desirable places to live in Boston.


**Part (h)**

```{r}
rm_over_7 <- subset(Boston, rm>7)
nrow(rm_over_7)  

rm("rm_over_7")
```


* There are 64 suburbs with more than 7 rooms per dwelling.

```{r}
rm_over_8 <- subset(Boston, rm>8)
nrow(rm_over_8)  

summary(rm_over_8)
```

* There are 13 suburbs with more than 7 rooms per dwelling

